Benchmarking on Rigetti architecture

### 2022-08-28 txt/manuscript.tex
i am done

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we will compare theoretical probability of correct 
discrimination between these measurements and the realization on real devices.
The theoretical probability can be calculated using Holevo-Helstrom theorem and 
Theorem 1 from~\cite{puchala2018strategies}. 
This probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme on real devices in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the proper input state, (see sec.~\ref{sec:example_discriminator}) 
and final measurement (see sec.~\ref{sec_example_final_measurement}).

@@

\begin{proposition}
Consider the problem of discrimination between von Neumann measurements $\PP_U$ 
and $\PP_\1$, $U = H\diag(1, e^{i \phi}) H^\dagger $ and $\phi \in [0, 
2\pi)$.  The optimal input state has the form
\begin{equation}
\ket{\psi_{AB}} = \frac{1}{\sqrt{2}} |\Id_2 \rangle \rangle.
\end{equation}
\end{proposition}

@@

	\begin{equation}
	\begin{split}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond & =  \max_{\rho \in \Omega_d} \| \left( \Id \otimes \sqrt{\rho} \right) J(\PP_{U} - \PP_{\Id} )  \left( \Id \otimes \sqrt{\rho} \right)  \|_1   \\ & =  \max_{\rho \in \Omega_d} \| \left( \Id \otimes \sqrt{\rho} \right) \sum_{i=0}^{d-1} \proj{i} \otimes \left( \proj{u_i} - \proj{i} \right)^\top  \|_1  \\ & = \max_{\rho \in \Omega_d} \| \sum_{i=0}^{d-1} \proj{i} \otimes \sqrt{\rho}  \left( \proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \|_1  \\ & = 
	 \max_{\rho \in \Omega_d} \| \sum_{i=0}^{d-1}\sqrt{\rho}  \left( \proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \|_1
	\end{split}
	\end{equation}
	We can prove that for all $\alpha, \beta \ge 0 $, and unit vectors $\ket{x}, \ket{y}$ the following equation holds 
	\begin{equation}
	\| \alpha \proj{x} - \beta\proj{y} \|_1 = \sqrt{(\alpha + \beta)^2 - 4\alpha \beta |\braket{x}{y}|^2}.
	\end{equation}
	By taking $\ket{x} = \frac{\sqrt{\rho} \ket{\bar{u_i}}}{\| \sqrt{\rho} \ket{\bar{u_i}} \|}$ and $ \ket{y} = \frac{\sqrt{\rho} \ket{i}}{\|\sqrt{\rho} \ket{i} \|}$ we obtain 
	\begin{equation}
		\| \PP_{U} - \PP_{\Id}  \|_\diamond  = \max_{\rho \in \Omega_d} \sum_{i=0}^{d-1} \sqrt{\left( \bra{u_i} \rho \ket{u_i} + \bra{i} \rho \ket{i }\right)^2 - 4 | \bra{u_i} \rho \ket{i} |^2}.
	\end{equation}
	\todo[inline]{dokonczyc}

	Let us consider  $\rho =   \frac{1}{2}  	\left(\begin{array}{cc}1&0\\0&1\end{array}\right)  $,   we obtain
	\begin{equation}
	\begin{split}
	||\mathcal{P}_U - \mathcal{P}_{\1}||_\diamond &= \sum_{i=0}^1  
	\sqrt{\left(\bra{u_i}\rho\ket{u_i} + \bra{i} \rho \ket{i} \right)^2 - 
		4|\bra{i}\rho\ket{u_i}|^2} = \\&  \sum_{i=0}^1  \sqrt{4 \bra{i}\rho 
		\ket{i}^2 - 4 \left| \frac{\lambda_1 \bra{i}\rho_1 \ket{i} + 
			\lambda_2\bra{i} \rho_2\ket{i}}{2}\right|^2} =\\&  \sum_{i=0}^1  \sqrt{4 
		\bra{i}\rho \ket{i}^2 - 4 \bra{i} \rho \ket{i}^2\left| \frac{\lambda_1 + 
			\lambda_2}{2}\right|^2} = \\&  \sum_{i=0}^1 2 \bra{i} \rho \ket{i} \sqrt{1 
		-\left| \frac{\lambda_1 + \lambda_2}{2}\right|^2 } = \\& 2 \sqrt{1 -\left| 
		\frac{\lambda_1 + \lambda_2}{2}\right|^2 } = 
	2 \sqrt{1 -\left| \frac{1+e^{i \phi}}{2}\right|^2 } = |1-e^{i \phi }|. 
	\end{split}
	\end{equation}
	From the definition of Choi matrix $J(\Phi)$ of $\Phi$ we have
	\begin{equation}
\| (\Id\otimes \sqrt{\rho} ) J(\Phi)  (\Id\otimes \sqrt{\rho} ) \|_1  =  \| (\Phi \otimes \Id) \left(  | \sqrt{\rho}^\top \rangle  \rangle \langle \langle \sqrt{\rho}^\top | \right) \|_1.
	\end{equation}
	So we obtain that \begin{equation}
	\ket{\psi_{AB}} =   | \sqrt{\rho}^\top \rangle  \rangle = \frac{1}{\sqrt{2} } | \Id_2 \rangle \rangle, 
	\end{equation}
	which completes the proof.
\end{proof}

@@

\begin{proposition}
The  final measurement has the form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{0} V_1^\dagger  \\ 
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{1} V_1^\dagger,
\end{split}
\end{equation}
where for each $\phi \in \mathbb{R}$,  the controlled unitaries $V_0$ and $V_1$ 
have the form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right) 
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
\end{proposition}
\todo[inline]{add proof of the proposition}

From ~\cite{puchala2018strategies}(Proposition 4) and Lemma~\ref{lemma:rho} we 
have 
\begin{equation}
\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
\rangle \rangle. 
\end{equation}
From Holevo-Helstrom theorem  we constrain a measurement $\mu$.  
Let us define \begin{equation}
X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
\otimes \Id_2 \right)(\proj{\psi})
\end{equation}
where $\ket{\psi}$ is defined in Remark~\ref{remark:discriminator}. From 
Hahn-Jordan decomposition let \begin{equation}
X = P - Q
\end{equation}
where $P, Q \ge 0 $. Observe, that $P $ and $Q$ are block-diagonal. 
Let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $im(P)$ and $im(Q)$, 
respectively. Then  $\Pi_P$ and $\Pi_Q$ have the following forms
\begin{equation}
\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
\end{equation}
and 
\begin{equation}
\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right) 
\end{equation}
Hence, we define $V_0$ such that
\begin{equation}
\begin{split}
V_0 \ket{x_p} = \ket{0} \\ 
V_0 \ket{x_q} = \ket{1}
\end{split}
\end{equation}
and $V_1$ such that
\begin{equation}
\begin{split}
V_1 \ket{y_p} = \ket{0} \\ 
V_1 \ket{y_q} = \ket{1}
\end{split}
\end{equation}
The explicit form of $V_0$ and $V_1$ we can see in \texttt{one-qubit.nb}.
Finally, we obtain the measurement $\mu$ of the form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{0} V_1^\dagger  \\ 
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{1} V_1^\dagger  
\end{split}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

\paragraph{Construction of $V_0$ and $V_1$}
	For each $\phi \in \mathbb{R}$,  the controlled unitary $V_0$ and $V_1$ 
	have the following form
	\begin{equation}
	V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
	\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
	\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
	\end{equation}

		\begin{equation}
	V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
	\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} 
	\right) &  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
	\end{equation}

\todo[inline]{obwody }
\todo[inline]{TU WYNIKI Z RIGETTIEGO}

>>>>

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we present theoretical probability of correct 
discrimination between these measurements. Moreover, we create the optimal theoretical strategy of their discriminate. 
Such a probability can be calculated using Holevo-Helstrom theorem and 
Theorem 1 from~\cite{puchala2018strategies}. 
This probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the proper input state, (see sec.~\ref{sec:example_discriminator}) 
and final measurement (see sec.~\ref{sec_example_final_measurement}).

@@

\begin{proposition}\label{prop-discrim}
Consider the problem of discrimination between von Neumann measurements $\PP_U$ 
and $\PP_\1$, $U = H\diag(1, e^{i \phi}) H^\dagger $ and $\phi \in [0, 
2\pi)$.  The optimal input state has the form
\begin{equation}
\ket{\psi_{AB}} = \frac{1}{\sqrt{2}} |\Id_2 \rangle \rangle.
\end{equation}
\end{proposition}

@@

	\begin{equation}
	\begin{split}
	\| \PP_{U} - \PP_{\Id}  \|_\diamond & =  \max_{\rho \in \Omega_2} \| \left( \Id \otimes \sqrt{\rho} \right) J(\PP_{U} - \PP_{\Id} )  \left( \Id \otimes \sqrt{\rho} \right)  \|_1   \\ & =  \max_{\rho \in \Omega_2} \| \left( \Id \otimes \sqrt{\rho} \right) \sum_{i=0}^{1} \proj{i} \otimes \left( \proj{u_i} - \proj{i} \right)^\top  \|_1  \\ & = \max_{\rho \in \Omega_2} \| \sum_{i=0}^{1} \proj{i} \otimes \sqrt{\rho}  \left( \proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \|_1  \\ & = 
	 \max_{\rho \in \Omega_2} \| \sum_{i=0}^{1}\sqrt{\rho}  \left( \proj{u_i} - \proj{i} \right)^\top \sqrt{\rho}  \|_1
	\end{split}
	\end{equation}
	We can prove that for all $\alpha, \beta \ge 0 $, and unit vectors $\ket{x}, \ket{y}$ the following equation holds 
	\begin{equation}
	\| \alpha \proj{x} - \beta\proj{y} \|_1 = \sqrt{(\alpha + \beta)^2 - 4\alpha \beta |\braket{x}{y}|^2}.
	\end{equation}
	By taking $\ket{x} = \frac{\sqrt{\rho} \ket{\bar{u_i}}}{\| \sqrt{\rho} \ket{\bar{u_i}} \|}$ and $ \ket{y} = \frac{\sqrt{\rho} \ket{i}}{\|\sqrt{\rho} \ket{i} \|}$ we obtain 
	\begin{equation}
		\| \PP_{U} - \PP_{\Id}  \|_\diamond  = \max_{\rho \in \Omega_2} \sum_{i=0}^{1} \sqrt{\left( \bra{u_i} \rho \ket{u_i} + \bra{i} \rho \ket{i }\right)^2 - 4 | \bra{u_i} \rho \ket{i} |^2}.
	\end{equation}
	\todo[inline]{dokonczyc}
	Let us take  $\rho_0 =   \frac{1}{2}  	\left(\begin{array}{cc}1&0\\0&1\end{array}\right)  $,   we obtain
	\begin{equation}
	\begin{split}
	||\mathcal{P}_U - \mathcal{P}_{\1}||_\diamond &= \sum_{i=0}^1  
	\sqrt{\left(\bra{u_i}\rho_0\ket{u_i} + \bra{i} \rho_0 \ket{i} \right)^2 - 
		4|\bra{i}\rho_0\ket{u_i}|^2} = \\& 
	\sum_{i=0}^1  \sqrt{ \braket{i}{i}^2 -  \left| \bra{i}  U \ket{i }\right|^2} = \\ & 
	 \sum_{i=0}^1  \sqrt{\braket{i}{i}^2 -  \left| 1 \cdot \bra{i} \proj{u_1} \ket{i} + 
			e^{i \phi} \cdot\bra{i}  \proj{u_2}\ket{i}\right|^2} =\\&  \sum_{i=0}^1  \sqrt{ 
		\braket{i}{i}^2 -\braket{i}{i}^2\left| \frac{1 + 
			e^{i \phi}}{2}\right|^2} = \\&  \sum_{i=0}^1\braket{i}{i} \sqrt{1 
		-\left| \frac{1+ e^{i \phi}}{2}\right|^2 } = 
	2 \sqrt{1 -\left| \frac{1+e^{i \phi}}{2}\right|^2 } = \frac{|1-e^{i \phi }|}{2}. 
	\end{split}
	\end{equation}
	Due to that, we know  
	that $| \sqrt{\rho}^{\top} \rangle \rangle$ is the discriminator of the problem of discrimination between 
	$\PP_{\Id} $ and $\PP_U$, where 
	$ U =  H \diag(1, e^{i \phi}) H^\dagger$ (from definition of diamond norm ).

	\begin{equation}
		\| \PP_{U} - \PP_{\Id}  \|_\diamond  =  \| \left( \Id \otimes \sqrt{\rho} \right) J(\PP_{U} - \PP_{\Id} )  \left( \Id \otimes \sqrt{\rho} \right)  \|_1  
=  \left\| ( (\PP_{U} - \PP_\Id) \otimes \Id) \left(  | \sqrt{\rho}^\top \rangle  \rangle \langle \langle \sqrt{\rho}^\top | \right) \right\|_1.
	\end{equation}
	So we obtain that \begin{equation}
	\ket{\psi_{AB}} =   | \sqrt{\rho}^\top \rangle  \rangle = \frac{1}{\sqrt{2} } | \Id_2 \rangle \rangle, 
	\end{equation}
	which completes the proof.
\end{proof}

@@

\begin{proposition}
	Consider the problem of discrimination between von Neumann measurements $\PP_U$ 
	and $\PP_\1$, $U = H\diag(1, e^{i \phi}) H^\dagger $ and $\phi \in [0, 
	2\pi)$.  
The  final measurement has the form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{0} V_1^\dagger  \\ 
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{1} V_1^\dagger,
\end{split}
\end{equation}
where for each $\phi \in \mathbb{R}$,  the controlled unitaries $V_0$ and $V_1$ 
have the form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} \right) 
&  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
\end{proposition}

\begin{proof}
From Proposition~\ref{prop-discrim} we have the form of discriminator given by
\begin{equation}
\ket{\psi} = | \sqrt{\rho}^\top \rangle \rangle = \frac{1}{\sqrt{2}} |\Id_2 
\rangle \rangle. 
\end{equation}
From Holevo-Helstrom theorem  we constrain a measurement $\mu$.  
Let us define \begin{equation}
X  = \left( \PP_U \otimes \Id_2 \right)(\proj{\psi}) -  \left( \PP_\Id 
\otimes \Id_2 \right)(\proj{\psi}).
\end{equation} From 
We consider the Hahn-Jordan decomposition, let \begin{equation}
X = P - Q
\end{equation}
where $P, Q \ge 0 $. Observe, that $P $ and $Q$ are block-diagonal. 
So, let us define projectors $\Pi_P$ and $\Pi_Q$ onto  $\text{im}(P)$ and $\text{im}(Q)$, 
respectively. Then  $\Pi_P$ and $\Pi_Q$ have the following forms
\begin{equation}
\Pi_P = \left(\begin{array}{cc}\proj{x_p}&0\\0&\proj{y_p}\end{array}\right) 
\end{equation}
and 
\begin{equation}
\Pi_Q = \left(\begin{array}{cc}\proj{x_q}&0\\0&\proj{y_q}\end{array}\right) 
\end{equation}
Hence, we define $V_0$ such that
\begin{equation}
\begin{cases} V_0 \ket{x_p} = \ket{0} \\ V_0 \ket{x_q} = \ket{1} \end{cases}
\end{equation}
and $V_1$ such that
\begin{equation}
\begin{cases} 
V_1 \ket{y_p} = \ket{0} \\ 
V_1 \ket{y_q} = \ket{1}
\end{cases} 
\end{equation}
For the discrimination task between $\PP_{U}$ and $\PP_{\Id}$, where $U = H \diag(1, e^{i \phi })H$, the explicit form of $V_0$ and $V_1$ is given by following (see also \texttt{one-qubit.nb}).
The unitary $V_0$ 
has the following form
\begin{equation}
V_0 = \left(\begin{array}{cc}i \sin\left( \frac{\pi - \phi}{4} \right)&-i 
\cos\left( \frac{\pi - \phi}{4} \right)\\ \cos\left( \frac{\pi - 
	\phi}{4}\right)& \sin\left( \frac{\pi - \phi}{4} \right)\end{array}\right),
\end{equation}
whereas the unitary  $V_1$  has the form
\begin{equation}
V_1 = \left(\begin{array}{cc}-i \cos\left(\frac{\pi - \phi}{4}\right) &i 
\sin\left( \frac{\pi - \phi}{4}\right)\\\sin\left( \frac{\pi - \phi}{4} 
\right) &  \cos\left( \frac{\pi - \phi}{4} \right) \end{array}\right).
\end{equation}
for all $\phi \in [0,2\pi)$. 
Finally, we obtain the measurement $\mu$ of the form
\begin{equation}
\begin{split}
\mu(0) = \proj{0} \otimes V_0 \proj{0} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{0} V_1^\dagger  \\ 
\mu(1) = \proj{0} \otimes V_0 \proj{1} V_0^\dagger +  \proj{1} \otimes V_1 
\proj{1} V_1^\dagger  
\end{split}
\end{equation}

\section{Implementation}
In this section we will compare theoretical probability of correct 
discrimination between these measurements and the realization on real devices.
\todo[inline]{TU WYNIKI PRAKTYTCZNE }
\todo[inline]{tu rozklad i obwody}
